PhD Title \\ PhD Thesis

### 2024-07-31 Latex_files/Thesis/chapters/chapter 2/main.tex
corrections made to chapters 2 and 4 of thesis doc

Optical tweezing operates on the principle that light carries both
linear and angular momentum, which can be transferred to other
objects. If light is reflected or refracted by a medium part of the
light's momentum is transferred to the medium itself. While at large
scales this effect is insignificant in most cases, at micron scales
the forces imparted by individual photons begin to become significant
enough to influence the trajectory of individual particles. Ashkin's 
work into optical tweezing found that micron sized entities could be 
spatially locked by aiming a laser with a Gaussian intensity profile 
directly upward at a the cell while being suspended in a liquid medium 
\cite{Ashkin1970}.

This is demonstrated in fig.~\ref{fig:working_principle} where we have 
a simple sphere trapped by a circularly symmetrical beam with the intensity
falling off from the centre in a normal distribution (commonly referred to
as a Gaussian beam). If we consider two rays of light hitting our sphere 
we see that while in the centre each ray is refracted in an equal opposite 
direction to one another and so the net force imparted is 0. When displaced 
in the x-y plane the net force is unbalanced and points towards the centre 
of the trap; this principle can be generalised to 3 degrees of freedom if we 
consider a focused beam instead of a para-axial beam, typically the trap will be
weaker along the beam axis compared to the transverse trap strength.

@@

Proper understanding of optical tweezing requires an understanding of
how the trapped particle interacts with the trapping laser. From a 
electromagnetism perspective the laser creates a spatially and 
temporally coherent electric field that scatters light off of a 
trapped particle.  The laws governing electric and magnetic fields 
are summarised most succinctly via the Maxwell equations. The 
differential forms of which are given below \cite{Jackson_1975}:
\begin{align}
  \nabla \cdot \mathbf{E}
  &= \frac{\rho_ v}{\epsilon_0}
  \\
  \nabla \cdot \mathbf{B}
  &= 0
  \\
  \nabla \times \mathbf{E}
  &= \frac{\partial \mathbf{B}}{\partial t}
  \\
  \nabla \times \mathbf{B}
  &= {\bf j} +\frac{\partial \mathbf{E}}{\partial t}    
\end{align}
where $\rho_v$ is the free charge density, ${\bf j}_0$ is the charge
current, and $\epsilon_0$ is the permittivity of free space.

These 4 equations describe how the electric and magnetic fields behave
and are related to one another at a microscopic level. Any discussion 
of optical trapping is underpinned by the fact that in all cases the 
Maxwell Equations must be satisfied at the macroscopic level, where 
one must account for the medium's interactions with the EM field. 
Macroscopic forms of the maxwell equations are given below \cite{Jackson_1975}:
\begin{align}
	\nabla \cdot \mathbf{D}
	&= \rho_f
	\\
	\nabla \cdot \mathbf{B}
	&= 0
	\\
	\nabla \times \mathbf{E}
	&= \frac{\partial \mathbf{B}}{\partial t}
	\\
	\nabla \times \mathbf{H}
	&= \mathbf{J}_f +\frac{\partial \mathbf{D}}{\partial t}  
\end{align}
Where $\rho_f$ and $\mathbf{J}_f$ are the charge density and current 
caused by the presence of free charges in the medium, and $\mathbf{D}$ 
and $\mathbf{H}$ are the displacement and magnetizing fields respectively. 
The latter two are defined by:
\begin{align}
	\nonumber
	\mathbf{D} &= \epsilon_0\mathbf{E}+\mathbf{P} 
	\\ 
	\nonumber
	\mathbf{H} &= \frac{1}{\mu_0}\mathbf{B}-\mathbf{M}
\end{align}

@@

The Ray-Optics model is the simplest to understand, the theory models 
the laser as a collection of individual 'rays' that propagate and are 
refracted according to Snell's Law. Based on the change in direction 
momentum is transferred to the target particle; with rays closest to 
the centre of the beam having greater intensity than those rays at 
the very edge of the beam. Consider a particle struck by two rays in 
a Gaussian beam (see Fig below), one coming close to the centre, and 
the other ray coming from the edge. As each ray is refracted by the 
target sphere a force is imparted onto it, the total force imparted 
is given by:
\begin{align}
	F_i = Q_i\frac{\Delta n P_i}{c}
\end{align}

where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference
in refractive indices between the solution and the target particle,
and $P_i$ is the power of the individual ray. The net force can be 
subdivided into its gradient and scattering components, where the latter
directs the particle to the centre of the beam, and the latter acts on 
the particle in the direction of beam propagation. For a beam with a
Gaussian intensity distribution $P_i$ will fall off as you move from
the centre of the beam.

@@

The Lorenz Mie theory provides an exact solution to the maxwell equations 
for the scattering caused by an isotropic sphere. The theory describes the 
scattered wave given off by a dielectric sphere when incident by a plane 
wave as a summation of partial spherical waves. For any spherical wave the 
vector fields must solve the Helmholtz wave equation given by:
\begin{align}
	\nabla^2\mathbf{E} +k^2\mathbf{E} = 0
\end{align}

Where $k$ is the wave number of the electromagnetic radiation ($k = 2\pi/\lambda$)
This combined with the constraints of Maxwell's Equations leaves 
very few exact solutions apart from spherical or planar waves, 
both of which can be converted between with relative ease. For 
example, a plane wave electric field can be expanded into spherical 
harmonics and likewise any spherical wave can be described as 
combination of plane waves offset from one another. For a single 
plane wave the incident, internal, and scattered fields are given as: 
\begin{align}
	\label{eq:incident}
	{\bf E}_{\rm inc}({\bf r})
	&=
    E_0 \sum_{nm}^\infty \left[
    a_{nm}{\bf M}_{nm}^{(1)}({\bf r})
    + b_{nm}{\bf N}_{nm}^{(1)}({\bf r})\right]
  \\
  \label{eq:internal}
  {\bf E}_{\rm int}({\bf r})
  &=
    E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left[
    -id_{nm}{\bf N}_{nm}^{(1)}({\bf r}) + c_{nm}{\bf M}_{nm}^{(1)}({\bf r})
    \right]
  \\
  \label{eq:scattered}
  {\bf E}_{\rm scat}({\bf r})
  &=
    E_0 \sum_{n=1}^\infty  i^n\frac{2n+1}{n(n+1)}\left[
      -i p_{nm}{\bf N}_{nm}^{(3)}({\bf r})+q_{nm}{\bf M}_{nm}^{(3)}({\bf r})
    \right] 
    \\
	\mathbf{E}_{tot}({\bf r}) &= 
	\begin{cases}
		{\bf E}_{\rm inc}({\bf r}) + {\bf E}_{\rm scat}({\bf r}) & \text{if \textbf{r} outside}
		\\ 
		{\bf E}_{\rm int}({\bf r}) & \text{if \textbf{r} inside}
	\end{cases}
\end{align}

where $a_{nm}$, $b_{nm}$, $c_{nm}$, $d_{nm}$, $p_{nm}$, and $q_{nm}$
are the expansion coefficients of each of the fields, and $M^{(1)}_{nm}$ 
and $N^{(1)}_{nm}$ are the vector spherical harmonics. For the incident
field computing its expansion coefficients is possible via analytical
methods and are completely dependent on the beam conditions imposed by
the user. However, computing the expansion coefficients of the
internal and scattered fields is far more tedious depending on the
shape of the target and what properties of the scattered field are
desired --- see sec.~\ref{sec:scattering}.  From an optical
trapping perspective the force exerted by a focused electric field can
be found by computing the Maxwell stress tensor which only requires
the total magnitude of the incident and scattered fields, essentially
computing the momentum contained in the incident beam and how much of
it has been transferred to the target particle.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Rayleigh Regime}

The Rayleigh approximation is the limiting approximation for describing 
a particles motion in an electromagnetic field who's wavelength is several 
times greater than the particle's size. The underlying theory is that a 
dielectric sphere can be treated as a dipole while in the presence of the 
electromagnetic field; in which case the scattering force is given simply 
by the scattering of the induced dipole, and the gradient force is due 
to the Lorentz force \cite{Gordon1973}. The gradient forces in the 
principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} 
in MKS units as a restorative rectangular force field:
\begin{align}
  F_{grad,x}
  &=-\hat{x} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \\
  F_{grad,y}
  &=-\hat{y} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \\
  F_{grad,z}
  &=-\hat{z} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2}
    \nonumber \times I(r)
  \\ 
  & \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
  \\
  \text{where:}
  \nonumber
  \\
	I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
	\exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
\end{align}

Where $m$ is the relative refractive index ($n_1/n_2$), $\omega_0$ is 
the beam waist, and $a$ is the radius of the particle. Scattering force 
however is dependent on the effective scattering cross sectional area. 
\begin{align}
  F_{\rm scat}
  &= \hat{z} \left(\frac{n_2}{2}\right) C_{pr} I(r) \\
  \text{where:} \nonumber \\
  C_{pr} &= \frac{8}{3} \pi (ka)^4 a^2 \left(\frac{m^2-1}{m^2+2}\right)^2
\end{align}

The Rayleigh regime allows for easy computation of the gradient and 
scattering forces due to the assumption that the particle is a point 
dipole, so much so that higher order scattering problems can be 
simplified by subdividing the particle into discrete dipoles (see 
Sec~\ref{sec:scattering}). However as the target particle gets larger 
this assumption fails to accurately describe the trapping force due 
to the complexity in the gradient field. For particle sizes close to 
the laser wavelength the scattered field is best described by Lorenz-Mie theory.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Scattering methods}
\label{sec:scattering}
There are several methods available to calculate the scattered field 
produced by a particle, while one can compute this from directly applying
Lorenz-Mie theory for complicated particle shapes and focused beams makes it 
tedious for some applications.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{T-matrix Method}
The $T$-matrix method was first developed by Peter Waterman with his
research into acoustic wave scattering \cite{Waterman1969}, this 
would later be extended to electromagnetic waves. Sometimes being 
referred to as the extended boundary condition method (ECBM), 
the method replaces the scatterer with a series of surface currents
over the targets surface. These currents are chosen so that the
electric field outside is identical to the original problem \cite{Wriedt1998}. 
This choice of currents negates the need to compute the internal fields
which reduces the scattering problem to a system of linear equations 
relating the incident beam coefficients to the scattered beam coefficients
the relationship between each can be summarised as:
\begin{align}
	\label{eq:t_matrix}
	\begin{pmatrix}
		q_{mn} \\
		p_{mn} 
	\end{pmatrix}
	= \bold{T} 
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
	= \begin{pmatrix}
		T_{11} \ T_{12} \\
		T_{21} \ T_{22}
	\end{pmatrix}
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
\end{align}

Where $q_{mn}$, and $p_{mn}$ are the scattering beam coefficients, and
$a_{mn}$, and $b_{mn}$ are the coefficients of the incident beam. For 
this PhD in particular we utilise an extension of the ECBM known as 
multi-sphere $T$-matrix method (\textit{mstm}). Developed by Mackowski
\cite{Mackowski2011} the computational code computes the scattered field 
from each sphere within the target cluster and the incident field itself. 
The scattering between spheres slowly converges to a final result but modern 
codes truncate the calculations to an upper number of iterations (denoted by
$N_{max}$) to fit a desired accuracy and computational time.

@@

Where $\mathbf{\bar{G}}$ is the Greens dyadic function of free space, which
defines the impulse response between two separate dipoles; and $\chi$ is the 
susceptibility of the medium, which describes the degree of polarisation of
the medium in the presence of an electric field. With the internal field 
calculated the scattered field can be computed, one of the primary advantages
of DDA over the T-matrix method is that the the composition of the target can 
be changed freely. When comparing different computational scattering methods,
the ECBM was found to be better suited for simulating the scattering of symmetric
targets as the ECBM can use the targets symmetry to speed up calculations \cite{Wriedt1998},
however when dealing with inhomogeneous media DDA is more efficient compared to
ECBM.

@@

The Langevin model can be extrapolated to describe the diffusive behaviour 
of an overall system, but for this project we can instead consider the 
behaviour of some particle with a diffusion tensor $D$ suspended in a 
viscous fluid and spacially localised by an optical potential with trap 
strength $\kappa$. Assuming the only external force acting on our particle 
is the laser the net force should be exactly equal to force of the stochastic
collisions due to the fluids thermal energy. If we focus our analysis when 
the particle is stably trapped and assume that the trap is harmonic we can 
model the trapping force as a Hookean spring ($F(t) \approx \kappa x(t)$). 
The full Langevin equation for an optically trapped particle is therefore given as:
\begin{align}
	\label{eq:langevin}
	m\frac{\delta^2x(t)}{\delta t^2} + \gamma_0 \frac{\delta x(t)}{\delta t} + \kappa_x x(t) = \sqrt{2k_BT}\eta(t)
\end{align}

\eqref{eq:langevin} provides an accurate description of strongly trapped 
particles, however the analytical solution of the Langevin equation requires 
integration of the white noise term making it difficult to simulate the 
trajectory of a given particle \cite{Volpe2013}. Instead it is often far 
easier to solve the equation numerically and apply use the analytical 
solution to calibrate and extract information about the particle and fluid, 
and how the two interact with the optical trap.

@@

Where $R_n(x(t))$ is the remainder term between the Taylor expansion to 
term n and the actual expression. If we limit our approach to the first 
derivative only we find that for sufficiently small values of $R_1$ the 
velocity and acceleration can be approximated as:
\begin{align}
	x'(t) &\approx \frac{x(t+\Delta t)-x(t)}{\Delta t}
	\\
	x^{''}(t) &\approx \frac{x'(t+\Delta t)-x'(t)}{\Delta t} = \frac{x(t)-2x(t+\Delta t)+x(t+2\Delta t)}{\Delta t^2}
\end{align}

By reversing the time step (i.e. use $-\Delta t$) to approximate the \\
velocity and acceleration based on the previous time steps we can 
discretize the position by taking finitely small  time steps (i.e. 
$x(t) = x_i,\ x(t-\Delta t) = x_{i-1}$). The same cannot be done for 
white noise as no information is known about $W(t)$ at any time. We 
can instead say that the velocity of a Brownian particle should 
approximate our white noise as a random walk, where at each new time 
step the position changes randomly within a given range. Constricting 
the variance to $\sqrt{2D}/\Delta t$ allows us to represent the white 
noise using the finite-difference approach as:
\begin{align}
	m\frac{x_i-2x_{i-1}+x_{i-2}}{\Delta t^2} = -\gamma\frac{x_i-x_{i-1}}{\Delta t}+\sqrt{2k_BT\gamma}\frac{w_i}{\sqrt{\Delta t}}
\end{align}

Where $w_i$ is a random real number between -1 and 1, we can say that 
it is normally distributed around 0 for simulation purposes. We can 
rearrange this for $x_i$ to approximate the Brownian motion of a 
particle (setting $x_0=0$), where the characteristic time is $\tau = m/\gamma$. 
Now in the case of an optical trap the restoration time scale is 
given by $\tau_{OT}=\kappa_x/\gamma$ which for strongly trapped particles 
is far greater than the characteristic time, therefore for simulation 
purposes we can neglect the particle's inertia which allows us to 
write the motion of an optically trapped particle as:
\begin{align}
	\label{eq:sim_langevin}
	x_i = x_{i-1} - \tau_{OT}\ x_{i-1}\Delta t + \sqrt{2D\Delta t}\ w_i
\end{align}

>>>>

Optical tweezing operates on the principle that light carries both
linear and angular momentum, which can be transferred to other
objects. If light is reflected or refracted by a medium part of the
light's momentum is transferred to the medium itself. While at large
scales this effect is insignificant in most cases, at micron scales
the forces imparted by individual photons begin to become significant
enough to influence the trajectory of individual particles. Ashkin's 
work into optical tweezing found that micron sized entities could be 
spatially locked by aiming a laser with a Gaussian intensity profile 
directly upward at a particle while being suspended in a liquid 
medium \cite{Ashkin1970}.

This is demonstrated in fig.~\ref{fig:working_principle} where we have 
a simple sphere trapped by a circularly symmetrical beam with the intensity
falling off from the centre in a normal distribution (commonly referred to
as a Gaussian beam). If we consider two rays of light hitting our sphere 
we see that while in the centre each ray is refracted in an equal opposite 
direction to one another and so the net force imparted is 0. When displaced 
in the x-y plane the net force is unbalanced and points towards the centre 
of the trap. This principle can be generalised to 3 degrees of freedom if we 
consider a focused beam instead of a para-axial beam, typically the trap will be
weaker along the beam axis compared to the transverse trap strength.

@@

Proper understanding of optical tweezing requires an understanding of
how the trapped particle interacts with the trapping laser. From an 
electromagnetism perspective the laser creates a spatially and 
temporally coherent electric field that scatters light off of a 
trapped particle.  The laws governing electric and magnetic fields 
are summarised most succinctly via the Maxwell equations. The 
differential forms of which are given below \cite{Jackson_1975}:
\begin{align}
  \nabla \cdot \mathbf{E}
  &= \frac{\rho_ v}{\epsilon_0}
  \\
  \nabla \cdot \mathbf{B}
  &= 0
  \\
  \nabla \times \mathbf{E}
  &= \frac{\partial \mathbf{B}}{\partial t}
  \\
  \nabla \times \mathbf{B}
  &= {\bf j} +\frac{\partial \mathbf{E}}{\partial t}    
\end{align}
where $\rho_v$ is the free charge density, ${\bf j}_0$ is the charge
current, and $\epsilon_0$ is the permittivity of free space.

These four equations describe how the electric and magnetic fields behave
and relate to one another at a microscopic level. Any discussion 
of optical trapping is underpinned by the fact that in all cases the 
Maxwell equations must be satisfied at the macroscopic level, where 
one must account for the medium's interactions with the EM field. 
Macroscopic forms of the Maxwell equations are given below \cite{Jackson_1975}:
\begin{align}
	\nabla \cdot \mathbf{D}
	&= \rho_f
	\\
	\nabla \cdot \mathbf{B}
	&= 0
	\\
	\nabla \times \mathbf{E}
	&= \frac{\partial \mathbf{B}}{\partial t}
	\\
	\nabla \times \mathbf{H}
	&= \mathbf{J}_f +\frac{\partial \mathbf{D}}{\partial t}  
\end{align}
Where $\rho_f$ and $\mathbf{J}_f$ are the charge density and current 
caused by the presence of free charges in the medium, and $\mathbf{D}$ 
and $\mathbf{H}$ are the displacement and magnetising fields respectively. 
The latter two are defined by:
\begin{align}
	\nonumber
	\mathbf{D} &= \epsilon_0\mathbf{E}+\mathbf{P} 
	\\ 
	\nonumber
	\mathbf{H} &= \frac{1}{\mu_0}\mathbf{B}-\mathbf{M}
\end{align}

@@

The Ray-Optics model is the simplest to understand, this theory models 
the laser as a collection of individual 'rays' that propagate and are 
refracted according to Snell's Law. Based on the change in direction 
momentum is transferred to the target particle; with rays closest to 
the centre of the beam having greater intensity than those rays at 
the very edge of the beam. Consider a particle struck by two rays in 
a Gaussian beam (see Fig below), one coming close to the centre, and 
the other ray coming from the edge. As each ray is refracted by the 
target sphere a force is imparted onto it, the total force imparted 
is given by:
\begin{align}
	F_i = Q_i\frac{\Delta n P_i}{c}
\end{align}

where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference
in refractive indices between the solution and the target particle,
and $P_i$ is the power of the individual ray. The net force can be 
subdivided into its gradient and scattering components, where the latter
directs the particle to the centre of the beam, and the latter acts on 
the particle in the direction of beam propagation. For a beam with a
Gaussian intensity distribution $P_i$ will fall off as you move from
the centre of the beam. The ray optics model is ideal when dealing 
with larger particles whose diameter far exceed the wavelength of light
being scattered. While it can be useful in predicting the forces 
experienced by said particle's it does not fully capture the behaviour
of light when considering interference between different rays.

@@

The Lorenz-Mie theory provides an exact solution to the Maxwell equations 
for the scattering caused by an isotropic sphere. This theory describes the 
scattered wave given off by a dielectric sphere when incident by a plane 
wave as a summation of partial spherical waves. For any spherical wave the 
vector fields must solve the Helmholtz wave equation given by:
\begin{align}
	\nabla^2\mathbf{E} +k^2\mathbf{E} = 0
\end{align}

Where $k$ is the wave number of the electromagnetic radiation 
($k = 2\pi/\lambda$). This combined with the constraints of 
Maxwell's equations leaves very few exact solutions apart from 
spherical or planar waves, both of which can be converted between 
with relative ease. For example, a plane wave electric field can 
be expanded into spherical harmonics and likewise any spherical 
wave can be described as a combination of plane waves offset from 
one another. For a single plane wave the incident, internal, and 
scattered fields are given as: 
\begin{align}
	\label{eq:incident}
	{\bf E}_{\rm inc}({\bf r})
	&=
    E_0 \sum_{nm}^\infty \left[
    a_{nm}{\bf M}_{nm}^{(1)}({\bf r})
    + b_{nm}{\bf N}_{nm}^{(1)}({\bf r})\right]
  \\
  \label{eq:internal}
  {\bf E}_{\rm int}({\bf r})
  &=
    E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left[
    -id_{nm}{\bf N}_{nm}^{(1)}({\bf r}) + c_{nm}{\bf M}_{nm}^{(1)}({\bf r})
    \right]
  \\
  \label{eq:scattered}
  {\bf E}_{\rm scat}({\bf r})
  &=
    E_0 \sum_{n=1}^\infty  i^n\frac{2n+1}{n(n+1)}\left[
      -i p_{nm}{\bf N}_{nm}^{(3)}({\bf r})+q_{nm}{\bf M}_{nm}^{(3)}({\bf r})
    \right] 
    \\
	\mathbf{E}_{tot}({\bf r}) &= 
	\begin{cases}
		{\bf E}_{\rm inc}({\bf r}) + {\bf E}_{\rm scat}({\bf r}) & \text{if \textbf{r} outside}
		\\ 
		{\bf E}_{\rm int}({\bf r}) & \text{if \textbf{r} inside}
	\end{cases}
\end{align}

where $a_{nm}$, $b_{nm}$, $c_{nm}$, $d_{nm}$, $p_{nm}$, and $q_{nm}$
are the expansion coefficients of each of the fields, and $M^{(1)}_{nm}$ 
and $N^{(1)}_{nm}$ are the vector spherical harmonics. For the incident
field computing its expansion coefficients is possible via analytical
methods and are completely dependent on the beam conditions imposed by
the user. However, computing the expansion coefficients of the
internal and scattered fields is far more tedious depending on the
shape of the target and what properties of the scattered field are
desired --- see sec.~\ref{sec:scattering}.  From an optical
trapping perspective the force exerted by a focused electric field can
be found by computing the Maxwell stress tensor which only requires
the total magnitude of the incident and scattered fields. By computing 
the momentum contained in the incident and scattered fields we can measure
the total momentum transferred to the target particle.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Rayleigh Scattering}

The Rayleigh approximation is the limiting approximation for describing 
a particles motion in an electromagnetic field whose wavelength is several 
times greater than the particle's size. The underlying theory is that a 
dielectric sphere can be treated as a dipole while in the presence of the 
electromagnetic field; in which case the scattering force is given simply 
by the scattering of the induced dipole, and the gradient force is due 
to the Lorentz force \cite{Gordon1973}. The gradient forces in the 
principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} 
in MKS units as a restorative rectangular force field:
\begin{align}
  F_{grad,x}
  &=-\hat{x} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \\
  F_{grad,y}
  &=-\hat{y} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \\
  F_{grad,z}
  &=-\hat{z} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2}
    \nonumber \times I(r)
  \\ 
  & \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
  \\
  \text{where:}
  \nonumber
  \\
	I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
	\exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
\end{align}

Where $m$ is the relative refractive index ($n_1/n_2$), $\omega_0$ is 
the beam waist, and $a$ is the radius of the particle. The scattering force 
however, is dependent on the effective scattering cross-sectional area. 
\begin{align}
  F_{\rm scat}
  &= \hat{z} \left(\frac{n_2}{2}\right) C_{pr} I(r) \\
  \text{where:} \nonumber \\
  C_{pr} &= \frac{8}{3} \pi (ka)^4 a^2 \left(\frac{m^2-1}{m^2+2}\right)^2
\end{align}

The Rayleigh regime allows for easy computation of the gradient and 
scattering forces due to the assumption that the particle is a point 
dipole. So much so that higher order scattering problems can be 
simplified by subdividing the particle into discrete dipoles (see 
Sec~\ref{sec:scattering}). However, as the target particle gets larger 
this assumption fails to accurately describe the trapping force due 
to the complexity in the gradient field. For particle sizes close to 
the laser wavelength the scattered field is best described by Lorenz-Mie 
theory.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Scattering methods}
\label{sec:scattering}
There are several methods available to calculate the scattered field 
produced by a particle. While one can compute this by directly applying
Lorenz-Mie theory, describing the interactions between focused beams and 
complicated shapes can be computationally slow and memory intensive. As a
result, several alternative scattering approaches have been developed that
allow for faster computation times simplifying the scattering problem.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{T-matrix Method}
The $T$-matrix method was first developed by Peter Waterman with his
research into acoustic wave scattering \cite{Waterman1969}, this 
would later be extended to electromagnetic waves. Sometimes  
referred to as the extended boundary condition method (ECBM), 
the method replaces the scatterer with a series of surface currents
over the targets surface. These currents are chosen so that the
electromagnetic field outside is identical to the original problem 
\cite{Wriedt1998}. This choice of currents negates the need to 
compute the internal fields which reduces the scattering problem 
to a system of linear equations relating the incident beam coefficients 
to the scattered beam coefficients. The relationship between each can 
be summarised as:
\begin{align}
	\label{eq:t_matrix}
	\begin{pmatrix}
		q_{mn} \\
		p_{mn} 
	\end{pmatrix}
	= \bold{T} 
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
	= \begin{pmatrix}
		T_{11} \ T_{12} \\
		T_{21} \ T_{22}
	\end{pmatrix}
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
\end{align}

Where $q_{mn}$, and $p_{mn}$ are the scattering beam coefficients, and
$a_{mn}$, and $b_{mn}$ are the coefficients of the incident beam. For 
this project we utilised an extension of the ECBM known as 
multi-sphere $T$-matrix method (\textit{mstm}). Developed by Mackowski
\cite{Mackowski2011} the computational code computes the scattered field 
from each sphere within the target cluster and the incident field itself. 
The scattering between spheres slowly converges to a final result but modern 
codes truncate the calculations to an upper number of iterations (denoted by
$N_{max}$) to fit a desired accuracy and computational time.

@@

Where $\mathbf{\bar{G}}$ is the Greens dyadic function of free space, which
defines the impulse response between two separate dipoles; and $\chi$ is the 
susceptibility of the medium, which describes the degree of polarisation of
the medium in the presence of an electric field. With the internal field 
calculated, the scattered field can be computed, one of the primary advantages
of DDA over the T-matrix method is that the composition of the target can 
be changed freely. When comparing different computational scattering methods,
the ECBM was found to be better suited for simulating the scattering of symmetric
targets as the ECBM can use the target's symmetry to speed up calculations \cite{Wriedt1998}. But when dealing with inhomogeneous media DDA is 
more efficient compared to ECBM.

@@

The Langevin model can be extrapolated to describe the diffusive behaviour 
of an overall system, but for this project we can instead consider the 
behaviour of some particle with a diffusion tensor $D$ suspended in a 
viscous fluid and spacially localised by an optical potential with trap 
strength $\kappa$. Assuming the only external force acting on our particle 
is the laser, the net force should be exactly equal to force of the stochastic
collisions due to the fluids thermal energy. If we focus our analysis when 
the particle is stably trapped and assume that the trap is harmonic, we can 
model the trapping force as a Hookean spring ($F(t) \approx \kappa x(t)$). 
The full Langevin equation for an optically trapped particle is therefore given as:
\begin{align}
	\label{eq:langevin}
	m\frac{\delta^2x(t)}{\delta t^2} + \gamma_0 \frac{\delta x(t)}{\delta t} + \kappa_x x(t) = \sqrt{2k_BT}\eta(t)
\end{align}

\eqref{eq:langevin} provides an accurate description of strongly trapped 
particles. Despite this the analytical solution of the Langevin equation requires 
integration of the noise term making it difficult to simulate the 
trajectory of a given particle \cite{Volpe2013}. Instead, it is often far 
easier to solve the equation numerically and use the analytical 
solution to calibrate and extract information about the particle and fluid, 
and how the two interact with the optical trap.

@@

Where $R_n(x(t))$ is the remainder term between the Taylor expansion to 
term n and the actual expression. If we limit our approach to the first 
derivative only, we find that for sufficiently small values of $R_1$ the 
velocity and acceleration can be approximated as:
\begin{align}
	x'(t) &\approx \frac{x(t+\Delta t)-x(t)}{\Delta t}
	\\
	x^{''}(t) &\approx \frac{x'(t+\Delta t)-x'(t)}{\Delta t} = \frac{x(t)-2x(t+\Delta t)+x(t+2\Delta t)}{\Delta t^2}
\end{align}

By reversing the time step (i.e. use $-\Delta t$) to approximate the \\
velocity and acceleration based on the previous time steps, we can 
discretise the position by taking finitely small  time steps (i.e. 
$x(t) = x_i,\ x(t-\Delta t) = x_{i-1}$). The same cannot be done for 
noise, as no information is known about $W(t)$ at any time. We 
can instead say that the velocity of a Brownian particle should 
approximate our noise as a random walk, where at each new time 
step, the position changes randomly within a given range. Constricting 
the variance to $\sqrt{2D}/\Delta t$ allows us to represent the 
noise using the finite-difference approach as:
\begin{align}
	m\frac{x_i-2x_{i-1}+x_{i-2}}{\Delta t^2} = -\gamma\frac{x_i-x_{i-1}}{\Delta t}+\sqrt{2k_BT\gamma}\frac{w_i}{\sqrt{\Delta t}}
\end{align}

Where $w_i$ is a random real number between -1 and 1, we can say that 
it is normally distributed around 0 for simulation purposes. We can 
rearrange this for $x_i$ to approximate the Brownian motion of a 
particle (setting $x_0=0$), where the characteristic time is $\tau = m/\gamma$. 
Now in the case of an optical trap, the restoration time scale is 
given by $\tau_{OT}=\kappa_x/\gamma$ which for strongly trapped particles 
is far greater than the characteristic time. Therefore for simulation 
purposes, we can neglect the particle's inertia which allows the motion 
of an optically trapped particle to be written as:
\begin{align}
	\label{eq:sim_langevin}
	x_i = x_{i-1} - \tau_{OT}\ x_{i-1}\Delta t + \sqrt{2D\Delta t}\ w_i
\end{align}

This result can be generalised for a 3-dimensional description of an 
optically trapped particle, where each Cartesian direction has its 
own unique characteristic restoration time. We see from the result 
that trajectory is dependent on only a handful of factors, the trap 
stiffness $\kappa_x$, the fluid viscosity $\gamma$, and the thermal 
energy of the system $k_BT$ (with the latter two being related by 
Einstein's formulation of the diffusion coefficient $D = k_BT/\gamma$). 
Therefore, by calculating these parameters to a high degree of precision 
allows one to get a precise description of the forces experienced by a 
target particle, which in the past has been used for highly accurate 
force transduction \cite{BergSoerensen2004, Smith2003}.
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.67\linewidth]{finite_differences.png}
	\caption{Example trajectory created using Finite Differences method
			 for a $2\mu m$ diameter sphere. Trap stiffness's were estimated using \textit{ott} at $\kappa_x = \kappa_y = -100\ pN/\mu m$ and $\kappa_z = -25\ pN/\mu m$. The particle's motion can be localised around the shaded ellipsoid.}
\end{figure}
